---
title: "EDA"
author: "Felix Tran"
date: "October 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


# Import data
```{r data_import, cache = TRUE}
library(rnoaa)

weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
weather_df
```

Looking at the output, group_by() doesn't appear to change anything, but it
affects how other functions from tidyverse work. ungroup() removes whatever
grouping options you put.

n() is how many counts.

summarize() gives us a dataset with summary information we specify for every 
group. In this case, the number of observations for each group we define 
(name by month) (36 groups because of 3 names by 12 months).
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n())
```

count() works in addition to n()
```{r}
weather_df %>% 
  count(month)
```

# Grouped summaries
Can look at descriptions of lots of stuff
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n(),
            num_na = sum(is.na(tmax)),
            mean_tmax = mean(tmax, na.rm = T),
            median_prcp = median(prcp, na.rm = T),
            sd_tmin = sd(tmin, na.rm = T))
```

summarize() outputs a tibble, so we can apply all the ggplot/tidyverse stuff
to our summarized results too.

We can stack graphs of our original data and our summary information to
create more complex graphs.
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n(),
            num_na = sum(is.na(tmax)),
            mean_tmax = mean(tmax, na.rm = T),
            median_prcp = median(prcp, na.rm = T),
            sd_tmin = sd(tmin, na.rm = T)) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point()


month_df <- weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n(),
            num_na = sum(is.na(tmax)),
            mean_tmax = mean(tmax, na.rm = T),
            median_prcp = median(prcp, na.rm = T),
            sd_tmin = sd(tmin, na.rm = T))

ggplot(weather_df, aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  geom_line(data = month_df, aes(x = month, y = mean_tmax, color = name))
```


Using spread() at the end of our summarize() function reformats our summarized
data into a format more suited for human readibility.
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(mean_tmax = mean(tmax, na.rm = T)) %>% 
  spread(key = name, value = mean_tmax)
```


# Grouped mutate
mutate() will apply the mutate function within the specified groups from 
group_by().
```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(mean_tmax = mean(tmax, na.rm = T), centered_tmax = tmax - mean_tmax) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) +
  geom_point()
```


# Window functions
min_rank in dplyr package preferred.
We can rank the tmax values by location and month, then see the coldest days of
each month at each location.
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  mutate(tmax_rank = min_rank(tmax)) %>% 
  filter(tmax_rank == 1)
```

Add desc within min_rank() to rank in opposite direction (in this case the 
hottest days)
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  mutate(tmax_rank = min_rank(desc(tmax))) %>% 
  filter(tmax_rank == 1)
```

lag() function in dplyr gives the preceding value of the specified variable.
This means you have to order the data in the way you want before applying lag().
In this example we want the max temp of yesterday, so the data are in order by 
date before applying the lag() function.
This still respects the group() function so it'll only take the preceding value
if the preceding value is within the same group.
```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(lag_tmax = lag(tmax))
```

Now we can calculate how much the temperature changes day to day in each 
location. Then we summarize() to see how variability there is in daily 
temperature changes in each location.
```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(lag_tmax = lag(tmax),
         one_day_change = tmax - lag_tmax) %>% 
  summarize(sd_one_day_change = sd(one_day_change, na.rm = T))
```


Note: summarize() only works with functions that return a single-number summary.
