---
title: "Visualization 1"
author: "Felix Tran"
date: "September 27, 2018"
output: html_document
---

# Visualizing data
You will make lots of graphs. Most are just for you, but you will make 
lots of them. Start with basic graphs, then build upon them using ggplot
to get better sense of your data.

# Set up session by loading libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
```

# Create dataset
```{r data_import, cache = TRUE}
library(rnoaa)

weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

# Creating graphs

Making our first plot.

The first parameter is the dataset, then the aesthetics 
(the things you care about showing). Since we haven't given
the plot the geometries we want, it gives us a blank plot.
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```

Adding geom_point() gives us a scatterplot with the specified aesthetics.
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

Can be used as part of a pipeline if desired
```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Plots from ggplot can be saved in a variable
```{r}
weather_sp <- ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
weather_sp
```

# More plot options
Add color to scatterplot by using name as the aesthetic.
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name))
```

Add another geom.

geom_smooth() plots a smooth line through your data. Include option to 
not include the SE bands around the line for this case.

Notice that the geom_smooth line is a single color because the color mapping
is only within geom_point.
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) +
  geom_smooth(se = FALSE)
```

Adjust alpha parameter in geom_point to adjust transperancy of points
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.4) +
  geom_smooth(se = FALSE)
```

When the aesthetics are moved to ggplot, all geoms are affected. 

geom_smooth graphs multiple lines with colors matching the points because the 
aesthetics differntiates the data into 3 clusters based on the locations 
(Central Park, HA, WA)
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.4) +
  geom_smooth(se = FALSE)
```

# Facetting
Create separate graphs for each value of name
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.4) +
  geom_smooth(se = FALSE) +
  facet_grid(~ name)
```


# More interesting plots
Look at max temperature over the course of a year
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point()
```

Instead of points, we can graph lines
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_line()
```

Add another aes option to make points small/large depending on value of
precipitation (prcp)
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name, size = prcp)) + 
  geom_point()
```

More smoothing options
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name, size = prcp)) + 
  geom_point() +
  geom_smooth() +
  facet_grid(~ name)
```


# Univariate plots
## Histograms
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) +
  geom_histogram()
```

## Density plot
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) +
  geom_density(alpha = 0.5)
```

## Box plot
Box plot aes is y-axis for the value, not x-axis like 
histogram and density plots. x-axis aes will help with 
separate box plots
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) +
  geom_boxplot()
```

## Violin plot
Convey distribution info like box and density plot combined
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) +
  geom_violin()
```

## Putting ggridges to work to make ridge plots
```{r}
ggplot(weather_df, aes(x = tmax, y = name)) +
  geom_density_ridges()
```

# Saving or embedding ggplots
Save as pdf
```{r}
ridge_plot <- ggplot(weather_df, aes(x = tmax, y = name)) +
  geom_density_ridges()
ggsave('ridge_plot.pdf', ridge_plot, width = 8, height = 5)
```

Options like this give you control for sizing/aspect ratio stuff for 
saved plots
```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

# Last example
```{r}
ggplot(weather_df, aes(y = prcp, x = name)) + geom_boxplot()
```
